Order reduction and MILP models for certifying Robustness of DNNs globally.

### 2025-08-01 AAAI2026/Lipitschz.tex
related
global pass

\title{Order reduction and partial MILP models \\
	for certifying Robustness of DNNs globally.}
\date{}
\author{
	%Authors
	% All authors must be in the same font size and format.
	Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
	AAAI Style Contributions by Pater Patel Schneider,
	Sunil Issar,\\
	J. Scott Penberthy,
	George Ferguson,
	Hans Guesgen,
	Francisco Cruz\equalcontrib,
	Marc Pujol-Gonzalez\equalcontrib
}
\affiliations{
	%Afiliations
	\textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
	% If you have multiple authors and multiple affiliations
	% use superscripts in text and roman font to identify them.
	% For example,

@@

	\hfill

>>>>

\title{Order reduction and MILP models for certifying Robustness of DNNs globally.}
\date{}
\author{
	%Authors
	% All authors must be in the same font size and format.
	Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
	AAAI Style Contributions by Pater Patel Schneider,
	Sunil Issar,\\
	J. Scott Penberthy,
	George Ferguson,
	Hans Guesgen,
	Francisco Cruz\equalcontrib,
	Marc Pujol-Gonzalez\equalcontrib
}
\affiliations{
	%Afiliations
	\textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
	% If you have multiple authors and multiple affiliations
	% use superscripts in text and roman font to identify them.
	% For example,

@@

	\vspace{-0.5cm}
	\input{related}

\section{Conclusion}

In this paper, we introduced a novel "2v" MILP model for global verification question, which is 4 to 8 times more accurate than the classical MILP model,
and more accurate than the ITNE model of \cite{lipshitz}, see Table \ref{table.classical}. Using our "3v" or "1v" models abstracting away some features,
 even better bounds are reached (Tables 3, 4, 5). We also explained how to support $L_1$ perturbations efficiently on top of classical $L_\infty$ perturbations (Prop \ref{prop.l1}). To further lower the bounds by $\approx 8$ times, we resort to model order reduction, reducing the space of search closer to the space of the training dataset, without sacrificing DNN accuracy ($97\%$). This results into, to the best of our knowledge, the first time real-time robustness certification is reported for a DNN, with $86\%$ of robustness for a $L_1$-perturbation of $.5$ and $53\%$ of robustness for a $L_1$-perturbation of $1$. Still, global verification is only at its beginning, and combining different methods is necessary to scale to more complex DNNs.

### 2025-08-01 AAAI2026/experiments.tex
global pass

	MNIST is a classification benchmark, we thus report different bound $\beta^{\epsilon}_{i,j}$
	for an $L_1$-deformation $\leq .5$. The pipe considers a conjunction of $L_1 \leq 3.9$ and $L_\infty \leq 0.2$ perturbations, corresponding to physically pertinent dimensions, and 
	maximizes the sum of the difference in output of 10 selected points in the mesh between an input and its deformation.

@@

@@

\begin{table}[h!]
	\begin{tabular}{|l|c|c|c|}\hline\hline
		model &        Bound $\downarrow$ &  Sol. &      Worst-Case $\uparrow$ \\\hline \hline
	Classical, $50 \times 2$ &    $.320$ &  $.320$ & $.017$ 
    \\\hline
	ITNE, $50 \times 2$ &    $.042$ &  $.037$ & $.022$
	\\ \hline
    2v model, $50 \times 2$ &    {\bf .040} &  $.037$ &  $.018$ 
    \\\hline \hline
    Classical, $100 \times 2$ &  .186  &  $.022$ & $.022$ 
    \\\hline
	ITNE, $100 \times 2$ &    $.045$ &  $.023$ & {\bf .023} 
    \\\hline
	2v model, $100 \times 2$&     {\bf .042} &  $.023$ &   {\bf .023} 
    \\\hline \hline
	\end{tabular}
	\caption{Comparison of the classical encoding, ITNE and our "2v" model on the pipe system with a fixed timeout of 1000s, with either $50 \times 2$, 
    or the full $100 \times 2$ binary variables.}
    \label{table.classical}
\end{table}

@@

	\begin{table}[h!]
		\centering
	\begin{tabular}{||l||c|c|c||}\hline\hline
		model &        Bound $\downarrow$ &  Sol. &      Worst-Case $\uparrow$ \\\hline \hline
		1v, $500 \times 1$ & $15.67$ & $.620$ & $.012$ \\\hline 
		3v, $500 \times 3$ & $16.09$ & $.755$ & {\bf .207} \\\hline 
	    2v, $500 \times 2$ & {\bf 8.99} & nan & nan \\\hline\hline	 
	\end{tabular}
	\caption{Comparison of "1v", "3v" and "2v" models 
	to obtain bounds on $\beta^{.5}_{6,8}$ on the {\bf full dimension} MNIST DNN, 
	for timeouts of $14400$s, when all 500 ($\times 1$, $\times 2$, $\times 3$) variables are binary.}
	\label{table.mnist}
\end{table}

@@

\paragraph{Reduced space}

@@

	\end{tabular}
	\caption{Comparison of "1v", "3v" and "2v" models 
	to obtain bounds on $\beta^{.5}_{6,8}$ on the {\bf 20 dimension}  reduced order MNIST DNN, for timeouts of 14400s, where 
	all the %400, 475,  or 
	500 ($\times 1$, $\times 2$, $\times 3$) neurons use binary variables.}
	\label{table.reduced}
\end{table}

To remove improbable images and limit the space of search, 
we consider a PCA model order reduction: We settled on reduction to 20 dimensions, because the MNIST $100 \times 5$ DNN considered, once run on images obtained from projecting to the reduced order and projected back to the full dimensional space displays the same accuracy of $97$\% as the DNN on the original images. This means that considering the reduced 20 dimensional space
does not incur any loss in accuracy, which is the only thing which matters. On this reduced space, bounds obtained are much more precise, and one can concretly certify in real-time robustness of images.

\begin{table}[b!]
	\begin{tabular}{||l||c|c|c|c||}\hline\hline
		model &    $L_1\leq 0.5$ & $L_1\leq 1$ & $L_1\leq 1.5$ &  $L_1\leq 2$ \\\hline \hline
		1v, $500\times1$ & $80 \%$ & $32\%$ & $7\%$ & $0\%$ \\\hline
		3v, $500 \times 3$ & {\bf 86 \%} & {\bf 53\%} & {\bf 20\%} & {\bf 4\%} \\\hline
		2v, $500 \times 2$ & 84\% & 51\% & 16\% & {\bf 4\%} \\\hline \hline
	\end{tabular}
	\caption{Comparison of "1v", "3v" and "2v" models 
	in the percentage of images certified robust in real-time 
	using the computed $(\beta^{.5}_{i,j})_{i < j \leq 10}$  for different values of $L_1$-perturbations.}
    \label{table.cert}
\end{table}

@@

	For the pipe system, we compared in Table \ref{table.pipe} the different models "1v","2v","3v" to produce bounds on the sum of the difference of strain over 10 specific points of the mesh, for a physically relevant perturbation of the deformation. The bounds we found are quite accurate, with a best bound of ".0329" obtained by the "3v" model, slightly better than the bound $.0337$ found within the same time by the "2v" model, and better than the bound found $0.356$ by the "1v" model, although this bound has been found 70 times faster due to the simpler model. The certified lower bound is not so far, at $.245$, found by the fully accurate "2v" model when all the variables are binary. We did check that this worst-case found, displayed in Fig. \ref{fig5}, which is not too far from the actual worst case $<.0337$, is coherent wit the physical Finite Element system the DNN has been learnt from, hence this is not an hallucination due to the brittleness of the DNN learnt.

@@

	\begin{table}[h!]
	\begin{tabular}{||l||c|c|c|c||}\hline\hline
		model &        Bound$\downarrow$ &  Sol. &      Worst-Case$\uparrow$ &  Time(s) \\\hline \hline
		1v, $100 \times 1$ &     {\bf .0356} &  $.0356$ & $.0191$ &  1000 \\\hline
		3v, $100 \times 3$&     .0414 &  .0254 &  .0166 &  1000 \\\hline
		2v, $100 \times 2$&     .0418 &  .0229 &   {\bf .0229} &  1000 \\\hline \hline
		%3v, $97 \times 3$&      ?? &  ?? &  ?? & 14440 \\\hline
		3v, $100 \times 3$&      {\bf .0350} &  .0272 &  .0216 & 14440 \\\hline
		%2v, $97 \times 2$&     ?? &  ?? &   ?? & 14440 \\\hline
		2v, $100 \times 2$&     .0360 &  .0236 &    {\bf .0236} & 14440 \\\hline \hline
		3v, $100 \times 3$&     {\bf .0329} &  .0277 &  .0165 & 72000 \\\hline
		2v, $100 \times 2$&     .0337 &  .0245 &  {\bf .0245} & 72000 \\\hline\hline
	\end{tabular}
	\caption{Comparison of "1v", "3v" and "2v" models on the pipe system with a fixed timeout of 1000s, 14440s and 72126s, where 90 ($\times 2$, $\times 3$) or all 100 ($\times 1, \times 2,\times 3$) neurons use binary variables.}
	%L1 corresponds to $3.9$ or $4$, and results should be the sum of 10 pixels, so around 10 times higher values.}
	\label{table.pipe}
\end{table}

>>>>

	MNIST is a classification benchmark, we thus report different bound $\beta^{\epsilon}_{i,j}$
	for an $L_1$-deformation $\leq .5$. The pipe considers a conjunction of 
	$L_1 \leq 3.9$ and $L_\infty \leq .02$ perturbations, corresponding to physically pertinent dimensions, and maximizes the sum of the difference in output of 10 selected points in the mesh between an input and its deformation.

@@

\begin{table}[b!]
	\centering
	\begin{tabular}{|l|c|c|c|}\hline\hline
		model &        Bound $\downarrow$ &  Sol. &      Worst-Case $\uparrow$ \\\hline \hline
	Classical, $50 \times 2$ &    $.320$ &  $.320$ & $.017$ 
    \\\hline
	ITNE, $50 \times 2$ &    $.042$ &  $.037$ & $.022$
	\\ \hline
    2v model, $50 \times 2$ &    {\bf .040} &  $.037$ &  $.018$ 
    \\\hline \hline
    Classical, $100 \times 2$ &  .186  &  $.022$ & $.022$ 
    \\\hline
	ITNE, $100 \times 2$ &    $.045$ &  $.023$ & .023
    \\\hline
	2v model, $100 \times 2$&     {\bf .042} &  $.023$ &   .023 
    \\\hline \hline
	\end{tabular}
	\caption{Comparison of the classical encoding, ITNE and our "2v" model on the pipe system with a fixed timeout of 1000s, with either $50 \times 2$, 
    or the full $100 \times 2$ binary variables.}
    \label{table.classical}
\end{table}

@@

@@

	\begin{table}[h!]
		\centering
	\begin{tabular}{||l||c|c|c||}\hline\hline
		model &        Bound $\downarrow$ &  Sol. &      Worst-Case $\uparrow$ \\\hline \hline
		1v, $500 \times 1$ & $15.67$ & $.620$ & $.012$ \\\hline 
		3v, $500 \times 3$ & $16.09$ & $.755$ & {\bf .207} \\\hline 
	    2v, $500 \times 2$ & {\bf 8.99} & n/a & n/a \\\hline\hline	 
	\end{tabular}
	\caption{Bounds on $\beta^{.5}_{6,8}$ 
	obtained by the "1v", "3v" and "2v" models 
	on the {\bf full dimension} MNIST DNN, 
	for timeouts of $14400$s, when all 500 ($\times 1$, $\times 2$, $\times 3$) variables are binary.}
	\label{table.mnist}
\end{table}

@@

@@

	\end{tabular}
	\caption{Comparison of "1v", "3v" and "2v" models 
	to obtain bounds on $\beta^{.5}_{6,8}$ on the {\bf 20 dimension} reduced order MNIST DNN, for timeout of 14400s, where 
	all %400, 475,  or 
	500 ($\times 1$, $\times 2$, $\times 3$) neurons use binary variables.}
	\label{table.reduced}
\end{table}

\paragraph{Reduced Space}

\begin{table}[b!]
	\begin{tabular}{||l||c|c|c|c||}\hline\hline
		model &    $L_1\leq 0.5$ & $L_1\leq 1$ & $L_1\leq 1.5$ &  $L_1\leq 2$ \\\hline \hline
		1v, $500\times1$ & $80 \%$ & $32\%$ & $7\%$ & $0\%$ \\\hline
		3v, $500 \times 3$ & {\bf 86 \%} & {\bf 53\%} & {\bf 20\%} & {\bf 4\%} \\\hline
		2v, $500 \times 2$ & 84\% & 51\% & 16\% & {\bf 4\%} \\\hline \hline
	\end{tabular}
	\caption{Percentage of images certified robust in real-time 
	using the computed $(\beta^{.5}_{i,j})_{i < j \leq 10}$ 
	by the "1v", "3v" and "2v" models, for different values of $L_1$-perturbations.}
    \label{table.cert}
\end{table}

To remove improbable images and limit the space of search, 
we consider a PCA model order reduction: We reduced to 20 dimensions, because the MNIST $100 \times 5$ DNN considered, once run on images obtained from projecting to the 20 dimension space and projected back to the full dimension space displays the same accuracy of $97$\% as the DNN on the original images. This means that considering the reduced 20 dimensional space
does not incur any loss in accuracy, which is the only thing which matters. On this reduced space, bounds obtained are much more precise, and one can certify in real-time robustness of images.

@@

	For the pipe system, we compared in Table \ref{table.pipe} the different models "1v","2v","3v" to produce bounds on the sum of the difference of strain over 10 specific points of the mesh, for a physically relevant perturbation of the deformation. The bounds we found are quite accurate, with a best bound of ".0329" obtained by the "3v" model, slightly better than the bound $.0337$ found within the same time by the "2v" model, and better than the bound found $0.356$ by the "1v" model, although this bound has been found 70 times faster due to the simpler model. The certified lower bound is not so far, at $.245$, found by the fully accurate "2v" model when all the variables are binary. We did check that this worst-case found, displayed in Fig. \ref{fig5} and which is not too far from the actual worst case that is known to be $<.0329$, is coherent wit the physical Finite Element system the DNN has been learnt from, hence this is not an hallucination due to the brittleness of the DNN learnt.

@@

	\begin{table}[h!]
	\begin{tabular}{||l||c|c|c|c||}\hline\hline
		model &        Bound$\downarrow$ &  Sol. &      Worst-Case$\uparrow$ &  Time(s) \\\hline \hline
		1v, $100 \times 1$ &     {\bf .0356} &  $.0356$ & $.0191$ &  1000 \\\hline
		3v, $100 \times 3$&     .0414 &  .0254 &  .0166 &  1000 \\\hline
		2v, $100 \times 2$&     .0418 &  .0229 &   {\bf .0229} &  1000 \\\hline \hline
		%3v, $97 \times 3$&      ?? &  ?? &  ?? & 14440 \\\hline
		3v, $100 \times 3$&      {\bf .0350} &  .0272 &  .0216 & 14440 \\\hline
		%2v, $97 \times 2$&     ?? &  ?? &   ?? & 14440 \\\hline
		2v, $100 \times 2$&     .0360 &  .0236 &    {\bf .0236} & 14440 \\\hline \hline
		3v, $100 \times 3$&     {\bf .0329} &  .0277 &  .0165 & 72000 \\\hline
		2v, $100 \times 2$&     .0337 &  .0245 &  {\bf .0245} & 72000 \\\hline\hline
	\end{tabular}
	\caption{Comparison of "1v", "3v" and "2v" models on the pipe system with timeouts of 1000s, 14440s and 72000s, where all 100 ($\times 1, \times 2,\times 3$) neurons use binary variables.}
	%L1 corresponds to $3.9$ or $4$, and results should be the sum of 10 pixels, so around 10 times higher values.}
	\label{table.pipe}
\end{table}

### 2025-08-01 AAAI2026/global.tex
global pass

			\addplot[only marks, mark=*, mark size=2pt, blue] coordinates {(2,2)};
			\node[label={above:$(\gamma_i,\gamma_i)$}] at (axis cs: 1.4, 1.5) {};
		\end{axis}
	\end{tikzpicture}
\caption{The possible values of $\hat{y}_i=\ReLU(x_i)-\ReLU(x'_i)$ depending on $y_i = x_i-x'_i$ for the "1v" model.}
	\label{fig.1v}
\end{figure}

@@

    We can easily check that the LP relaxation of this MILP program is exactly 
    $\frac{y_i-\gamma_i}{2} \leq \hat{y}_i \leq \frac{y_i+\gamma_i}{2}$, 
    {\bf LIAO please do on Fig2: as depicted in broken lines on Fig.~\ref{fig.1v}}.

@@

	The "1v" model contains the same number of (binary) variables as the MILP model for local robustness, but it is not as accurate as the "3v" or "2v" models.

>>>>

			\addplot[only marks, mark=*, mark size=2pt, blue] coordinates {(2,2)};
			\node[label={above:$(\gamma_i,\gamma_i)$}] at (axis cs: 1.4, 1.5) {};
		\end{axis}
	\end{tikzpicture}
\caption{The possible values of $\hat{y}_i=\ReLU(x_i)-\ReLU(x'_i)$ depending on $y_i = x_i-x'_i$ for the "1v" model. {\bf LIAO please depict LP relaxation in broken lines}.
    }
	\label{fig.1v}
\end{figure}

@@

    We can easily check that the LP relaxation of this MILP program,
	depicted in broken lines on Fig.~\ref{fig.1v}, is exactly: 
    \begin{equation}
		\label{eq.lpr}
	  \frac{y_i-\gamma_i}{2} \leq \hat{y}_i \leq \frac{y_i+\gamma_i}{2}
	\end{equation}

	As the "1v" model is a projection of the "2v" and the "3v" models, 
	the LP relaxations of the "2v" and the "3v" models contain (\ref{eq.lpr}).
	Notice that (\ref{eq.lpr}) corresponds to Eq. (3) in \cite{lipshitz}, 
	which is added explicitly to the classical model in their ITNE model.

@@

	%The "1v" model contains the same number of (binary) variables as the MILP model for local robustness, but it is not as accurate as the "3v" or "2v" models.

### 2025-08-01 AAAI2026/introduction.tex
global pass

   \item  In terms of perturbations, we consider conjunctions of $L_\infty$ and $L_1$ norms, which allows to accurately describe perturbations. For instance, "each subpixel is perturbed by at most $\frac{20}{255}$ ($L_\infty$) and the sum of the absolute value of perturbations over all subpixels is at most $\frac{150}{255}$" ($L_1$ perturbation). While $L_1$ perturbations are not linear (because of the absolute values), reason for which it is seldom used, Prop.~\ref{prop3} shows how to use it as a perturbation in the MILP model without incurring any expansive binary variables (only cheap linear variables are necessary).

>>>>

   \item  In terms of perturbations, we consider conjunctions of $L_\infty$ and $L_1$ norms, which allows to accurately describe perturbations. For instance, "each subpixel is perturbed by at most $\frac{20}{255}$ ($L_\infty$) and the sum of the absolute value of perturbations over all subpixels is at most $\frac{150}{255}$" ($L_1$ perturbation). While $L_1$ perturbations are not linear (because of the absolute values), reason for which it is seldom used, Prop.~\ref{prop.l1} shows how to use it as a perturbation in the MILP model without incurring any expansive binary variables (only cheap linear variables are necessary).

### 2025-08-01 AAAI2026/related.tex
global pass

The most related work is Vhagar \cite{vhagar}, which also computes $\beta^\varepsilon_{i,j}$ bounds, but limited to $L_\infty$ perturbations (and variants). They do not however report actual 
real-time certified robust percentage, as we do in Table \ref{table.cert}. Last, their hyper-adversarial attack to find better lower bounds seems extremely efficient, and we will consider it in the future.

Technically, none of the MILP base work \cite{vhagar,ITNE,GROCET} 
consider $L_1$ bounds. Further, they use the classical MILP encoding from \cite{MILP}, and not our novel "2v" model, nor even our "1v" model. However, \cite{ITNE,GROCET} 
defines the {\em diff variables} $y,\hat{y}$, adding explicitely the linear relaxation $\frac{y_i-\gamma_i}{2} \leq \hat{y}_i \leq \frac{y_i+\gamma_i}{2}$ implied by our models, Eq. (3) from \cite{ITNE} corresponding essentially to it. \cite{vhagar} does things differently, adding per neuron constraints (depending on the perturbation) between the (binary) variables to simplify the MILP problem.

>>>>

The most related work is Vhagar \cite{vhagar}, which also computes $\beta^\varepsilon_{i,j}$ bounds, but limited to $L_\infty$ perturbations (and variants). They do not however report real-time certified robust percentage, as in Table \ref{table.cert}. Last, their hyper-adversarial attack to find better lower bounds seems very efficient, and we will consider it in the future.

Technically, none of the MILP base work \cite{vhagar,ITNE} 
consider $L_1$ bounds. Further, they use the classical MILP encoding from \cite{MILP}, and not our novel "2v" model, nor even our "1v" model. However, \cite{ITNE} 
defines the {\em diff variables} $y,\hat{y}$, adding explicitely the linear relaxation $\frac{y_i-\gamma_i}{2} \leq \hat{y}_i \leq \frac{y_i+\gamma_i}{2}$ implied by our models, Eq. (3) in \cite{ITNE} corresponding essentially to it. \cite{vhagar} does things differently, adding per neuron constraints (depending on the perturbation) between the (binary) variables to simplify the MILP problem.
